The zero--one laws of Kolmogorov and Hewitt--Savage in categorical probability

### 2019-11-04 categorical-zero-one.tex
Add references for the two zero-one laws

Very roughly, the Kolmogorov zero--one law states that, given an infinite collection of independent random variables, and an event which is in the $\sigma$-algebra generated by the variables, which is independent of any finite subset of variables, the event has probability one or zero.\improve{add references for the Kolmogorov and Hewitt--Savage 01 laws}

The Hewitt--Savage zero--one law is similar - here the variables are supposed to be independent and indetically distributed, and the event is supposed to be invariant under finite permutations of the variables. Again, under these assumptions, the event has probability one or zero.

@@

In $\FinStoch$, the deterministic morphisms $X \to Y$ are exactly the $\{0,1\}$-valued stochastic matrices. Since these can be identified with honest functions $X \to Y$, we conclude that $\FinStoch_\detc$ is equivalent to the category of finite sets. In $\Stoch$, the deterministic morphisms $f : X \to Y$ are exactly those Markov kernels for which $f(S|x) \in \{0,1\}$ for every $S \in \Sigma_Y$ and $x \in X$. In other words, as the term suggests, the deterministic morphisms are those which do not involve any randomness. In $\CRing_+\op$, the deterministic morphisms $R \to S$ are precisely those morphisms which are represented by additive\todo{should this be ``multiplicative''?}  unital maps $S \to R$, i.e.~the ring homomorphisms. Thus the subcategory of deterministic morphisms is exactly the opposite of the usual category of commutative rings.

>>>>

Very roughly, the Kolmogorov zero--one law states that, given an infinite collection of independent random variables, and an event which is in the $\sigma$-algebra generated by the variables, which is independent of any finite subset of variables, the event has probability one or zero.
It goes back to Kolmogorov's foundational monograph \cite{kolmogorovgerm} - although the statement appearing there is significantly different from the modern form.
For a more recent treatment of this theorem, see e.g. \cite[Theorem~2.37]{klemke}

The Hewitt--Savage zero--one law, due to Hewitt and Savage in \cite[Theorem~11.3]{hewitt_savage} is similar - here the variables are supposed to be independent and indetically distributed, and the event is supposed to be invariant under finite permutations of the variables. Again, under these assumptions, the event has probability one or zero.

@@

In $\FinStoch$, the deterministic morphisms $X \to Y$ are exactly the $\{0,1\}$-valued stochastic matrices. Since these can be identified with honest functions $X \to Y$, we conclude that $\FinStoch_\detc$ is equivalent to the category of finite sets. In $\Stoch$, the deterministic morphisms $f : X \to Y$ are exactly those Markov kernels for which $f(S|x) \in \{0,1\}$ for every $S \in \Sigma_Y$ and $x \in X$. In other words, as the term suggests, the deterministic morphisms are those which do not involve any randomness. In $\CRing_+\op$, the deterministic morphisms $R \to S$ are precisely those morphisms which are represented by multiplicative unital maps $S \to R$, i.e.~the ring homomorphisms. Thus the subcategory of deterministic morphisms is exactly the opposite of the usual category of commutative rings.
